Formal Verification of Real-Time Data Processing of the LHC Beam Loss Monitoring System: A Case Study

### 2011-03-28 fmics/paper.tex
Add last two theorems to main text, remove Figure 9
last theorem doesn't have proof yet

\begin{theorem}
For all values of $D$, $n$, and $x$, the $\mathsf{output}$ of slice $n$ at an update time $t$ satisfies
\begin{equation*}%\label{output_input_at_update_times}
%\forall{D\,n\,x\,t}.\,{0 < n}\land{\mathsf{update\_time\;n\;t}}\implies\\
\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)-1}\begin{cases}0 \qquad\qquad\qquad  t<m+\mathsf{delay\_sum}\;n\\D\;(t-m-\mathsf{delay\_sum}\;n)\quad\text{otherwise}\end{cases}
\end{equation*}
\end{theorem}

@@

In light of Theorem~(\ref{output_last_update}), which says a slice's output stays constant between update times for the slice, Theorem~(\ref{output_input_at_update_times}) characterises all outputs at all times.
Thus, the SRS structure's outputs are exact running sums of the input counts over windows whose sizes depend on the the width of the slice and position of taps.
However, the sums are delayed by the total delay across all previous slices (represented by $\mathsf{delay\_sum}$).

\begin{figure}
\caption{
A selection of theorems about our model proved using HOL4.
\label{fig:theorems}
}
\begin{align}
\label{update_time_last_update}
&\forall{n\,t}.\,\mathsf{update\_time}\;n\;(\mathsf{last\_update}\;n\;t)\\
%\label{last_update_thm}
%&\forall{n\,t}.\,\mathsf{last\_update}\;n\;t=t-t\operatorname{mod}\mathsf{delay}\;n\\
%SR_prev
%SR_first
\label{output_last_update}
&\forall{D\,n\,x\,t}.\,\mathsf{output}\;D\;n\;x\;t=\mathsf{output}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t)\\
\label{output_sum}
&\forall{D\,n\,x\,t}.\,{0<n}\implies\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\mathsf{SR}\;D\;n\;m\;t
\end{align}
%\begin{multline}
%\label{output_source_at_update_times}
%\forall{D\,n\,x\,t}.\,{0<n}\land{\mathsf{update\_time\;n\;t}}\implies\\
%\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\begin{cases}0&t\leq m\times(\mathsf{delay}\;n)\\\mathsf{source}\;D\;n\;0\;(t-m\times(\mathsf{delay\;n})-1)&\text{otherwise}\end{cases}
%\end{multline}
\begin{multline}
\label{output_prev_at_update_times}
\forall{D\,n\,x\,t}.\,{0<n}\land{\mathsf{update\_time}\;n\;t}\implies\\
\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\begin{cases}0&t < m\times\mathsf{delay}\;n+\mathsf{delay}\;n'\\\mathsf{output}\;D\;n'\;x'\;(t-m\times\mathsf{delay}\;n-\mathsf{delay}\;n')&\text{otherwise}\end{cases}\\
\qquad\text{where $(n',x')=\mathsf{input}\;n$}
\end{multline}
\begin{multline}\label{output_input_at_update_times}
\forall{D\,n\,x\,t}.\,{0 < n}\land{\mathsf{update\_time\;n\;t}}\implies\\\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)-1}\begin{cases}0&t<m+\mathsf{delay\_sum}\;n\\D\;(t-m-\mathsf{delay\_sum}\;n)&\text{otherwise}\end{cases}
\end{multline}
\begin{multline}\label{output_eq_exact}
\forall{D\,n\,x\,t}.\,{0 < n}\implies\\\mathsf{output}\;D\;n\;x\;t=\begin{cases}0&\mathsf{last\_update}\;n\;t<\mathsf{delay}\;n\\\mathsf{exact}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t+1-\mathsf{delay\_sum}\;n)\end{cases}
\end{multline}
%max_error
\end{figure}

>>>>

\begin{theorem}\label{output_input_at_update_times}
For all values of $D$, $n$, and $x$, the $\mathsf{output}$ of slice $n>0$ at an update time $t$ satisfies
\begin{equation*}%\label{output_input_at_update_times}
%\forall{D\,n\,x\,t}.\,{0 < n}\land{\mathsf{update\_time\;n\;t}}\implies\\
\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)-1}\begin{cases}0 \qquad\qquad\qquad  t<m+\mathsf{delay\_sum}\;n\\D\;(t-m-\mathsf{delay\_sum}\;n)\quad\text{otherwise}\end{cases}
\end{equation*}
\end{theorem}

@@

Since a slice's output stays constant between update times for the slice, Theorem~\ref{output_input_at_update_times} characterises all outputs at all times.
Thus, the SRS structure's outputs are exact running sums of the input counts over windows whose sizes depend on the the width of the slice and position of taps.
However, the sums are delayed by the total delay across all previous slices (represented by $\mathsf{delay\_sum}$).

The fact that the SRS calculates exact, but delayed, sums, is captured in the next theorem, which relates $\mathsf{output}\;D\;n\;x\;t$ to $\mathsf{exact}\;D\;n\;x$ at an earlier time.

\begin{theorem}\label{output_eq_exact}
For all values of $D$, $n$, and $x$, the $\mathsf{output}$ of slice $n>0$ at all times $t$ satisfies
\begin{equation*}
\mathsf{output}\;D\;n\;x\;t=\begin{cases}0\qquad\qquad\qquad\qquad\qquad\quad\mathsf{last\_update}\;n\;t+1<\mathsf{delay\_sum}\;n\\\mathsf{exact}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t+1-\mathsf{delay\_sum}\;n)\quad\text{otherwise}\end{cases}
\end{equation*}
\end{theorem}
\begin{proof}
Follows from Theorem~\ref{output_input_at_update_times}, using the definition of $\mathsf{exact}$, and the fact that $\mathsf{output}\;D\;n\;x\;t$ has the same value as at the last update time.\qed
\end{proof}

We have also proved a preliminary result bounding the error of $\mathsf{output}\;D\;n\;x\;t$ compared to $\mathsf{exact}\;D\;n\;x\;t$.
\begin{theorem}\label{max_error}
For all $k$ and $D$ satisfying $|D\;(t'+1)-D\;t'|\leq k$ at all times, the following holds for all slices $n>0$, tap positions $x$, and times $t$:
\begin{multline*}
t>(\mathsf{tap}\;n\;x+1)\times(\mathsf{delay}\;n)+\mathsf{delay}\;n+\mathsf{delay\_sum}\;n\implies\\
|\mathsf{output}\;D\;n\;x\;t-\mathsf{exact}\;D\;n\;x\;t|\leq(\mathsf{tap}\;n\;x+1)\times(\mathsf{delay\;n})\times k\times(\mathsf{delay}\;n+\mathsf{delay\_sum}\;n)
\end{multline*}
\end{theorem}
\begin{proof}
\end{proof}

%\begin{figure}
%\caption{
%A selection of theorems about our model proved using HOL4.
%\label{fig:theorems}
%}
%\begin{align}
%\label{update_time_last_update}
%&\forall{n\,t}.\,\mathsf{update\_time}\;n\;(\mathsf{last\_update}\;n\;t)\\
%%\label{last_update_thm}
%%&\forall{n\,t}.\,\mathsf{last\_update}\;n\;t=t-t\operatorname{mod}\mathsf{delay}\;n\\
%%SR_prev
%%SR_first
%\label{output_last_update}
%&\forall{D\,n\,x\,t}.\,\mathsf{output}\;D\;n\;x\;t=\mathsf{output}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t)\\
%\label{output_sum}
%&\forall{D\,n\,x\,t}.\,{0<n}\implies\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\mathsf{SR}\;D\;n\;m\;t
%\end{align}
%%\begin{multline}
%%\label{output_source_at_update_times}
%%\forall{D\,n\,x\,t}.\,{0<n}\land{\mathsf{update\_time\;n\;t}}\implies\\
%%\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\begin{cases}0&t\leq m\times(\mathsf{delay}\;n)\\\mathsf{source}\;D\;n\;0\;(t-m\times(\mathsf{delay\;n})-1)&\text{otherwise}\end{cases}
%%\end{multline}
%\begin{multline}
%\label{output_prev_at_update_times}
%\forall{D\,n\,x\,t}.\,{0<n}\land{\mathsf{update\_time}\;n\;t}\implies\\
%\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\begin{cases}0&t < m\times\mathsf{delay}\;n+\mathsf{delay}\;n'\\\mathsf{output}\;D\;n'\;x'\;(t-m\times\mathsf{delay}\;n-\mathsf{delay}\;n')&\text{otherwise}\end{cases}\\
%\qquad\text{where $(n',x')=\mathsf{input}\;n$}
%\end{multline}
%\begin{multline}\label{output_eq_exact}
%\forall{D\,n\,x\,t}.\,{0 < n}\implies\\\mathsf{output}\;D\;n\;x\;t=\begin{cases}0&\mathsf{last\_update}\;n\;t<\mathsf{delay}\;n\\\mathsf{exact}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t+1-\mathsf{delay\_sum}\;n)\end{cases}
%\end{multline}
%%max_error
%\end{figure}
